### 2017-03-27 paper/oopsla17-submission/gallery.tex
more gallery

\paragraph{Computing the Fixed Point of a Function}
% data/sp14/0941.ml
Finally, our students must write a |fixpoint| function that computes the
fixed point of a given function |f|, starting from an initial value |b|.
%
As a hint we first have them write a |wwhile| function that performs the
functional equivalent of a while-loop, repeatedly passing a function's
output back in until it receives a (boolean) signal to stop.
%
\begin{ecode}
  let rec wwhile (f, b) =
    match f b with
    | (x, false) -> x
    | (x, true)  -> wwhile (f, x)
\end{ecode}
\lstset{firstnumber=last}
%
The |fixpoint| function can then be written as a clever instantiation of
the arguments to |wwhile|.
%
\begin{ecode}
  let fixpoint (f, b) =
    let g = __let bb = f b in ___=(bb, (bb = b))=_ in
    wwhile (g, b)
\end{ecode}
\lstset{firstnumber=1}
%
Sadly, our student has forgotten that |g| should itself be a function.
%
As a result, she passes a pair of a \emph{pair} and a starting value to
|wwhile|, rather than a pair of a \emph{function} and a starting value.
%
\sherrloc deduces that the call to |wwhile| is likely correct (\ocaml
actually blames the use of |g|), but identifies the construction of the
pair inside |g| as the most likely culprit, while our Decision Tree
correctly identifies the \emph{definition} of |g| as the source of the
error.
\ES{TODO: WHY?}

\subsubsection{Failed Predictions}
\label{sec:failed-predictions}
Of course, our classifiers are sometimes wrong, we focus next on the
10\% of programs where \emph{none} of our Top-3 predictions are correct.

% data/sp14/0148.ml
\begin{ecode}
  let rec clone x n =
    let loop acc n =
      if n <= 0 then
        acc
      else
        clone ==([==_=x=_==] @ acc)== (n - 1) in
    loop [] n
\end{ecode}
\ES{TODO: our 2nd prediction matches \sherrloc and \ocaml (occurs check), correct fix is to replace recursive call to clone with loop}

% data/sp14/0234.ml
\begin{ecode}
  let pipe fs =
    let f a x z = _=z=___ a__ in
    let base y = y in
    List.fold_left f base fs
\end{ecode}
\ES{TODO: fix is to swap a and z, \sherrloc is correct, \ocaml blames use of f in fold-left..}

% data/sp14/0255.ml
\begin{ecode}
  let rec wwhile (f, b) =
    let check = f b in
    let (x, y) = (f, b) in
    if __y__ = false then
      x
    else
      wwhile (f, ==x==)
\end{ecode}
\ES{marking ``y'' seems totally bogus, why is it even part of the error
  slice? (other locations are ``f b'' and ``false'' though, whihc is not
  any better)}

>>>>

\paragraph{Computing the Fixed Point of a Function}
% data/sp14/0941.ml
Finally, our students must write a |fixpoint| function that computes the
fixed point of a given function |f|, starting from an initial value |b|.
%
As a hint we first have them write a |wwhile| function that performs the
functional equivalent of a \texttt{while}-loop, repeatedly passing a function's
output back in until it receives a (boolean) signal to stop.
%
\begin{ecode}
  let rec wwhile (f, b) =
    match f b with
    | (x, false) -> x
    | (x, true)  -> wwhile (f, x)
\end{ecode}
\lstset{firstnumber=last}
%
The |fixpoint| function can then be written as a clever instantiation of
the arguments to |wwhile|.
%
\begin{ecode}
  let fixpoint (f, b) =
    let g = __let bb = f b in ___=(bb, (bb = b))=_ in
    wwhile (g, b)
\end{ecode}
\lstset{firstnumber=1}
%
Sadly, our student has forgotten that |g| should itself be a function.
%
As a result, she passes a pair of a \emph{pair} and a starting value to
|wwhile|, rather than a pair of a \emph{function} and a starting value.
%
\sherrloc deduces that the call to |wwhile| is likely correct (\ocaml
actually blames the use of |g|), but identifies the construction of the
pair inside |g| as the most likely culprit, while our Decision Tree
correctly identifies the \emph{definition} of |g| as the source of the
error.
\ES{TODO: WHY?}

\subsubsection{Failed Predictions}
\label{sec:failed-predictions}
Of course, our classifiers are sometimes wrong, we focus next on the
$\sim$10\% of programs where \emph{none} of our Top-3 predictions are correct.

\paragraph{Cloning an Element}
% data/sp14/0148.ml
Consider the following implementation of the |clone| function from
before.
%
\begin{ecode}
  let rec clone x n =
    let loop acc n =
      if n <= 0 then
        acc
      else
        clone ==([==_=x=_==] @ acc)== (n - 1) in
    loop [] n
\end{ecode}
% \ES{TODO: our 2nd prediction matches \sherrloc and \ocaml (occurs check), correct fix is to replace recursive call to clone with loop}
%
The student has defined a helper function |loop| with an accumulator
|acc|, likely meant to call itself tail-recursively.
%
Unfortunately, she has called the top-level function |clone| rather than
|loop| in the |else| branch, this induces a cyclic constraint |'a = 'a list|
for the |x| argument to |clone|.

Our classifier incorrectly predicts that the use of |x| in the recursive
call is the most likely source of the error.
\ES{TOOD: WHY??}
%
Our second prediction coincides with \sherrloc (and \ocaml), blaming the
the first argument to |clone|.
%
This is also incorrect, but may be more helpful than our first
prediction --- if our student decides that she has certainly provided
the correct \emph{argument}, an alternative explanation is that
perhaps she has called the wrong \emph{function}.
%
Our final prediction is the entire body of |clone|, which is hardly
helpful.

\paragraph{Composing Functions}
% data/sp14/0234.ml
Next, let us consider the |pipe| function that composes a list of
functions, \ie given a list of functions |[f;g;h]|, |pipe| produces the
function |fun x -> f (g (h x))|.
%
\begin{ecode}
  let pipe fs =
    let f a x z = _=z=___ a__ in
    let base y = y in
    List.fold_left f base fs
\end{ecode}
%
Leaving aside issues of \emph{functional} correctness, the error in the
student's code is that she has swapped function and argument in the
application |z a|.
%
\sherrloc correctly blames the use of |z|, while our classifier
(incorrectly) blames the entire application (\ocaml blames the use of
|f| on line 4).
\ES{TODO: WHY??}
%

\paragraph{A Functional \texttt{while}-loop}
% data/sp14/0255.ml
Finally, consider the following attempt at the functional
\texttt{while}-loop from before.
%
\begin{ecode}
  let rec wwhile (f, b) =
    let check = f b in
    let (x, y) = (f, b) in
    if __y__ = false then
      x
    else
      wwhile (f, ==x==)
\end{ecode}
% ES: the let (x, y) = (f, b) is actually a match, but this seems easier to present..
Our student

\ES{marking ``y'' seems totally bogus, why is it even part of the error
  slice? (other locations are ``f b'' and ``false'' though, which is not
  any better)}
